---
title: "project 4"
author: "Dipesh Poudel"
date: "1/4/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Repeated K-Fold Cross Validation

## Reading the File
```{r}
library(haven)
bank_loan_df <- read_sav("P4_bankloan_5000_clients.sav")
```

## Changing the data type of variables
```{r}
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)
```



## Splitting the data into train and test set
```{r}
set.seed(1234)
library(caret)
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
train_data<-bank_loan_df[ind==1,]
test_data<-bank_loan_df[ind==2,]
```

## Setting Up the Train Control
```{r}
rep_cv_train_control<-trainControl(method = "repeatedcv",number = 10,repeats = 3)
```

## Logistic Regression With Repeated Cross Validation

### Training Logistic Regression Model
```{r warning=FALSE}
logistic_clf1<-train(defaulted_loan~.,
  data=train_data,
  method="glm",
  family="binomial",
  trControl=rep_cv_train_control
)
summary(logistic_clf1)
```
### Making the Prediction
```{r}
predicted_val_log1<-predict(logistic_clf1,newdata = test_data)
```

### Confusion Matrix for Evaluation
```{r}
confusionMatrix(predicted_val_log1,test_data$defaulted_loan)
```
## KNN Model with Repeated Cross validation
### Training KNN Model
```{r}
knn_clf1<-train(defaulted_loan~.,data = train_data,
               method="knn",
                 trControl=rep_cv_train_control
               )
```

### Getting the Result of the Model
```{r}
knn_clf1$result
```
### Confusion Matrix for Model Evaluation 
```{r}
predicted_val_knn1<-predict(knn_clf1,newdata = test_data)
```

```{r}
confusionMatrix(predicted_val_knn1,test_data$defaulted_loan)
```

## Naïve Bayes classifier

### Training the Model
```{r}
library(naivebayes)
nb_clf1<-train(defaulted_loan~.,
               data=train_data,
               method="naive_bayes",
               usepoisson = TRUE,
               trControl=rep_cv_train_control
               )
```

```{r}
summary(nb_clf1)
```

### Making Prediction on Test Data
```{r}
predicted_val_nb1<-predict(nb_clf1,newdata = test_data)
```

### Confusion Matrix for Model Evaluation

```{r}
confusionMatrix(predicted_val_nb1,test_data$defaulted_loan)
```

## Support Vector Machine (SVM) Model

### Training the Model
```{r}
svm_clf1<-train(defaulted_loan~.,
               data=train_data,
               method="svmLinear",
               trControl=rep_cv_train_control,
               )
```

```{r}
svm_clf1
```

### Making the Prediction for test data
```{r}
predicted_val_svm1<-predict(svm_clf1,newdata = test_data)
```

### Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_svm1,test_data$defaulted_loan)
```
## Decision Tree Model
```{r}
dtree_clf1<-train(defaulted_loan~.,
                 data = train_data,
                 method="rpart",
                 parms = list(split = "information"),
                 tuneLength=10,
                 trControl=rep_cv_train_control
                 )
```

```{r}
dtree_clf1
```

### Making the Prediction for test data
```{r}
predicted_val_dtree1<-predict(dtree_clf1,newdata = test_data)
```

### Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_dtree1,test_data$defaulted_loan)
```

## Artifical Neural Network (ANN) Model

### Training the Model
```{r}
ann_clf1 <- train(defaulted_loan ~ ., data = train_data, 
  method = "nnet",
  preProcess = c("center","scale"), 
  maxit = 250,    # Maximum number of iterations
  tuneGrid = data.frame(size = 1, decay = 0),
  # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
  metric = "Accuracy",
  trControl=rep_cv_train_control)
```

### Making the Predictions for Test data
```{r}
predicted_val_ann1<-predict(ann_clf1,newdata = test_data)
```

### Confusion Matrix for the Model Evaluation
```{r}
confusionMatrix(predicted_val_ann1,test_data$defaulted_loan)
```

