---
title: "Project 4"
author: "Dipesh Poudel"
date: "12/28/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(haven)
bank_loan_df <- read_sav("P4_bankloan_5000_clients.sav")
```

```{r}
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)
```

```{r}
str(bank_loan_df)
```

Use the attached SPSS data containing 5000 cases of bank clients and fit all the supervised classifiers models as described below. Copy all the outputs you will get at/from R/R Studio here and write interpretation for each of the outputs in this file.

Use "defaulted_loan" variable as dependent variable and use other variables as independent variables to predict by:

Fitting logistic regression classifier, KNN classifier, ANN-MLP classifier, Naïve Bayes classifier, SVM classifier and Decision Tree classifier with train/test validation sets, leave-one-out cross validation sets, 10-fold cross validation sets and 3 times repeated 10-fold cross validation sets for this problem and select the best predictive model and find the important variables (variable importance) with relative frequencies and its graph for that model

Fitting bagging, random forest and extreme gradient boosting tree models and select the best predictive model for train/test validation sets and find the variable importance with table and graph for that model

Who will get the loan as per the best models identified above? Why? Explain with justifications.

# Train Test Validation

```{r}
library(caret)
```

## Splitting the data into traning and testing set

```{r}
set.seed(1234)
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
train_data<-bank_loan_df[ind==1,]
test_data<-bank_loan_df[ind==2,]
```

## Logistic Regression With train/test Validation
### Training Logistic Regression Model
```{r warning=FALSE}
logistic_clf<-train(defaulted_loan~.,
  data=train_data,
  method="glm",
  family="binomial"
)
summary(logistic_clf)
```
### Making the Prediction
```{r}
predicted_val_log<-predict(logistic_clf,newdata = test_data)
```

### Confusion Matrix for Evaluation
```{r}
confusionMatrix(test_data$defaulted_loan, predicted_val_log)
```
## KNN Model with train/test validation
### Training KNN Model
```{r}
knn_clf<-train(defaulted_loan~.,data = train_data,
               method="knn",
               preProcess = c("center", "scale"),
               tuneLength = 10
               )
```

### Getting the Result of the Model
```{r}
knn_clf$result
```
### Confusion Matrix for Model Evaluation 
```{r}
predicted_val_knn<-predict(knn_clf,newdata = test_data)
confusionMatrix(test_data$defaulted_loan, predicted_val_knn)
```
## Naïve Bayes classifier
### Training the Naïve Bayes classifier
```{r}
library(e1071)
nb_clf<-naiveBayes(defaulted_loan~.,data=train_data)
```

### Summary of the Model
```{r}
(nb_clf)
```
### Making the Prediction in the test data
```{r}
predicted_val_nb<-predict(nb_clf,newdata = test_data)
```

### Confusion Matrix for Evaluation
```{r}
confusionMatrix(predicted_val_nb,test_data$defaulted_loan)
```
## Support Vector Machine (SVM) Model

### Training the Model
```{r}
svm_clf<-train(defaulted_loan~.,
               data=train_data,
               method="svmLinear"
               )
svm_clf
```
### Making the Prediction for test data
```{r}
predicted_val_svm<-predict(svm_clf,newdata = test_data)
```

### Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_svm,test_data$defaulted_loan)
```
## Decision Tree Model
```{r}
dtree_clf<-train(defaulted_loan~.,
                 data = train_data,
                 method="rpart",
                 parms = list(split = "information"),
                 tuneLength=10
                 )
dtree_clf
```
### Making the Prediction for test data
```{r}
predicted_val_dtree<-predict(dtree_clf,newdata = test_data)
```

### Confusion Matrix for Model Evaluation
```{r}
confusionMatrix(predicted_val_dtree,test_data$defaulted_loan)
```

## Artifical Neural Network (ANN) Model

### Training the Model
```{r}
ann_clf <- train(defaulted_loan ~ ., data = train_data, 
  method = "nnet",
  preProcess = c("center","scale"), 
  maxit = 250,    # Maximum number of iterations
  tuneGrid = data.frame(size = 1, decay = 0),
  # tuneGrid = data.frame(size = 0, decay = 0),skip=TRUE, # Technically, this is log-reg
  metric = "Accuracy")
```

### Making the Predictions for Test data
```{r}
predicted_val_ann<-predict(ann_clf,newdata = test_data)
```

### Confusion Matrix for the Model Evaluation
```{r}
confusionMatrix(predicted_val_ann,test_data$defaulted_loan)
```