---
title: "Project 4"
author: "Dipesh Poudel"
date: "1/5/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Fitting bagging, random forest and extreme gradient boosting tree models and select the best predictive model for train/test validation sets and find the variable importance with table and graph for that model 

# Bagging, Boosting and Random Forest
## Reading the File
```{r}
library(haven)
bank_loan_df <- read_sav("P4_bankloan_5000_clients.sav")
```

## Changing the data type of variables
```{r}
bank_loan_df$defaulted_loan<-as.factor(bank_loan_df$defaulted_loan)
bank_loan_df$education_level<-as.factor(bank_loan_df$education_level)
```



## Splitting the data into train and test set
```{r}
set.seed(1234)
library(caret)
ind<-sample(2,nrow(bank_loan_df),replace=T,prob = c(0.7,0.3))
train_data<-bank_loan_df[ind==1,]
test_data<-bank_loan_df[ind==2,]
```

## Bagging Model

### Training the Model
```{r}
library("ipred")
bag_dtree_clf<-bagging(defaulted_loan~.,
                       data = train_data,
                       coob=T
                       )
```

```{r}
print(bag_dtree_clf)
```
### Making the Prediction
```{r}
predicted_bag_tree<-predict(bag_dtree_clf,newdata = test_data)
```

```{r}
library(caret)
confusionMatrix(predicted_bag_tree,test_data$defaulted_loan)
```
## Random Forest Model

### Training the Model
```{r}
set.seed(1234)
library(randomForest)
rf_clf<-randomForest(defaulted_loan~.,
                       data = train_data)
```

```{r}
rf_clf
```
### Making the Prediction
```{r}
predicted_rf<-predict(rf_clf,newdata = test_data)
```

```{r}
confusionMatrix(predicted_rf,test_data$defaulted_loan)
```

## Extreme Gradient Boosting

### Training the Model
```{r}
xglm_clf<-train(defaulted_loan~.,
                data = train_data,
                method="xgbTree",
                verbose=F
                )
```

```{r}
xglm_clf
```
### Making the Prediction
```{r}
predicted_xgb<-predict(xglm_clf,newdata = test_data)
```

```{r}
confusionMatrix(predicted_xgb,test_data$defaulted_loan)
```
```{r}
summary(xglm_clf)
```
```{r}
```


```{r}
varImp(xglm_clf)
```


```{r}
plot(varImp(xglm_clf))
```

