---
title: "Project 1"
author: "Dipesh Poudel"
date: "10/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 1: Import, Tidy, Transform, Visualize and Summarize API data

## Loading the Data from API
1. Import this JSON API data "https://data.askbhunte.com/api/v1/covid" using "jsonlite" or similar package Save the imported data as "covidtbl" data frame in R (Hint: use "flatten = TRUE" after 'fromJSON' function availble in jsonlite package)

```{r}
library(jsonlite)
url<-'https://data.askbhunte.com/api/v1/covid'
covidtbl<-fromJSON(url,flatten=TRUE)
covidtbl<-subset(covidtbl,select = -c(relatedTo,point.type,point.coordinates))
print(summary(covidtbl))
```
## Checking For Duplicate
3. Check if there are duplicate cases in the data using "id" variable, remove duplicate cases, if found, using R base functions: duplicated or unique (read about it and use it carefully) 
```{r}
# Finding the Numbers of Unique ids 
print(length(unique(covidtbl$id)))
# Number of total number of rows 
print(nrow(covidtbl))
```
Since the number of unique ids and total number of columns are same we can say that there are no duplicates rows based on Id

## Cleaning and Analyzing 
4. Clean the "gender" variable and show the number and percentage of males and females in the data (Hint: you can use table function and prop.table functions)
### Finding the unique values in the gender variable
```{r}
unique(covidtbl$gender)
```
As we can see from the result above there are two issues. First we need to take care of missing values and second we need to make data of same case.

```{r}
# Finding Percentage of NULL Values
dim(covidtbl)
```

