---
title: "Project 1"
author: "Dipesh Poudel"
date: "10/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 1: Import, Tidy, Transform, Visualize and Summarize API data

## Loading the Data from API
1. Import this JSON API data "https://data.askbhunte.com/api/v1/covid" using "jsonlite" or similar package 

2. Save the imported data as "covidtbl" data frame in R (Hint: use "flatten = TRUE" after 'fromJSON' function availble in jsonlite package)

```{r}
library(jsonlite)
url<-'https://data.askbhunte.com/api/v1/covid'
covidtbl<-fromJSON(url,flatten=TRUE)
# Removing Some Columns that we do not need for Analysis 
covidtbl<-subset(covidtbl,select = -c(relatedTo,point.type,point.coordinates))
print(summary(covidtbl))
```
## Checking For Duplicate
4. Check if there are duplicate cases in the data using "id" variable, remove duplicate cases, if found, using R base functions: duplicated or unique (read about it and use it carefully) 
```{r}
# Finding the Numbers of Unique ids 
print(length(unique(covidtbl$id)))
# Number of total number of rows 
print(nrow(covidtbl))
```
Since the number of unique ids and total number of columns are same we can say that there are no duplicates rows based on Id

## Cleaning and Analyzing gender variable 
5. Clean the "gender" variable and show the number and percentage of males and females in the data (Hint: you can use table function and prop.table functions)
### Finding the unique values in the gender variable
```{r}
unique(covidtbl$gender)
```
### Repacing NA with unknown

```{r}
covidtbl$gender[is.na(covidtbl$gender)] <- 'unknown'
unique(covidtbl$gender)
```
### Making the values of same case
```{r}
covidtbl$gender<-tolower(covidtbl$gender)
```

### Finding the total number of male and female and percentage

```{r}
# Finding the Numbers of Male, Female and Unknown
ugender<-as.data.frame(table(covidtbl$gender))
ugender$percentage<-(ugender$Freq/sum(ugender$Freq))*100
ugender
```
From the table above we can see that 27.50% of reported cases were female and 72.41% of reported cases were male.
## Cleaning and Analuzing Age Variable
6. Clean the "age" variable and show the summary statistics of the age variable and interpret the results carefully (Search the online newspapers and clean the age data, if required)

### Dealing with the NA values


```{r}
summary(covidtbl$age)
```
From the summary above we can see that there two issues that we need to deal with. First one is the max age is 523 which is impossible and second one is we have Null Value. 
```{r}
library(dplyr)
filter(covidtbl,age>=100)
```
As we can see there is only one row that is causing the issue. We will remove this row.

```{r}
# Making a copy Just in case something goes wrong
covidtbl_rm<-covidtbl
# covidtbl_rm<-covidtbl_rm[!(covidtbl_rm$age>=200),]
covidtbl[covidtbl$id==532, "age"] <- median(covidtbl$age,na.rm=TRUE)
```


```{r}
hist(covidtbl_rm$age,main = "Histogram of Age After Dealing Outlier",xlab = "Age")
```
Based on the histogram we will use median to replace the missing value.
```{r}
covidtbl$age[is.na(covidtbl$age)]<-median(covidtbl$age,na.rm=TRUE)
```
```{r}
summary(covidtbl_rm$age)
hist(covidtbl_rm$age,main = "Histogram of Age",xlab = "Age")
```

```{r}
summary(covidtbl$age)
```

As we can see from the summary, we now do not have outlier and NA values in the age variable.

## Transforming the continous variable into catagorical
7. Transform cleaned age variable into broad age groups i.e. <15, 15-59, 60+ years, define it as factor variable and get number and percentage of this variable and interpret it carefully

### Creating a new categorical variable from age 

```{r}
covidtbl<-within(covidtbl,{
  age.cat<-NA
  age.cat[age<15]<-"<15"
  age.cat[age>=15 & age<60]<-'15-59'
  age.cat[age>=60]<-'60+'
})
```
### Converting the new variable into factor.
```{r}
covidtbl$age.cat <- factor(covidtbl$age.cat, levels = c("<15", "15-59", "60+"))
summary(covidtbl$age.cat)
```

```{r}
uage<-as.data.frame(table(covidtbl$age.cat))
uage$percentage<-(uage$Freq/sum(uage$Freq))*100
uage
```

As we can see that the about 99.5% of reported cases were in the age group of 15 to 59. It seems realistic as the covid in Nepal first broke out in the cluster of people who worked outside and country and returned after pandemic. Most test was also done in the same cluster.

## Finding difference between recovered and reported dates
8. Find the number of days between recovered and reported dates (define it as: diff1 variable), clean it if required, and get the summary statistics of this variable and interpret it carefully

### Converting the variables into date type

```{r}
covidtbl$reportedOn<-as.Date(covidtbl$reportedOn)
covidtbl$recoveredOn<-as.Date(covidtbl$recoveredOn)
```
### Getting the difference in days
```{r}
covidtbl$diff1<-as.numeric(difftime(covidtbl$recoveredOn, covidtbl$reportedOn, units = "days"))
```

### Getting of summary
```{r}
summary(covidtbl$diff1)
```
From the summary we can see that it took 27 days on average for people to recover from covid. The maximum number of days was 102 and minimum was 0.

## Finding the Difference between reported date and death date
9. Find the number of days between deaths and reported dates (define it as: diff2 variable), clean it if required, and get the summary statistics of this variable and interpret it carefully
### Converting the deathOn and reportedOn to date
```{r}
covidtbl$deathOn<-as.Date(covidtbl$deathOn)
```
### Getting the difference in days
```{r}
covidtbl$diff2<-as.numeric(difftime(covidtbl$deathOn, covidtbl$reportedOn, units = "days"))
```
### Geting the Summary of New variable
```{r}
summary(covidtbl$diff2)
```
The Minimum value is -8 which is kind of odd. Lets check the rows with negative difference.

```{r}
library(dplyr)
filter(covidtbl,diff2<0)
```
There are 29 rows where the value of diff1 is negative. This is possible since there were few cases in which the cause of death was determined to be covid after the person was dead and no test was done beforehand.


## Appropriate Measure of Central Tendency and Dispersion
10. Which measures of central tendency and dispersion is most appropriate for the age, diff1 and diff2 variables? Why? Justify the use the appropriate measure of central tendency and dispersion by creating and interpreting the histogram and boxplot for the age, diff1 and diff2 variables and their careful interpretations

### Histogram for Age,diff1 and diff2
```{r}
hist(covidtbl$age,main = "Histogram of Age", xlab = "Age")
hist(covidtbl$diff1,main = "Histogram of diff1",xlab="Diff1")
hist(covidtbl$diff2,main = "Histogram of diff2", xlab = "Diff2")
```
### Boxplot for Age,diff1 and diff2
```{r}
boxplot(covidtbl$age,main="Boxplot of Age")
boxplot(covidtbl$diff1, main="Boxplot of diff1")
boxplot(covidtbl$diff2, main="Boxplot of diff2")
```
Analyzing the graphs, we can see that the data is skewed for age, diff1 and diff2 so we have to use median for central tendency and Inter Quartile Range for IQR.

## Analysis of current state variable
11. Show the number and percentage of the "current state" variable and interpret it carefully

```{r}
unique(covidtbl$currentState)
# Replacing the Missing value with Unknown
covidtbl$currentState[is.na(covidtbl$currentState)] <- 'unknown'
unique(covidtbl$currentState)
csper<-as.data.frame(table(covidtbl$currentState))
csper$percentage<-(csper$Freq/sum(csper$Freq))*100
csper
```

From the result we can see that out of 77817 observations (reported case of covid) 72.5% recovered from the covid and 0.6% people died are other cases are still active as of date of collection the data.  

## Analyzing the isReinfected variable
12. Show the number and percentage of the "isReinfected" variable, what percentage of cases were re-infected in Nepal at the given time period in the database? Was it realistic?

```{r}
summary(covidtbl$isReinfected)
ureinfected<-as.data.frame(table(covidtbl$isReinfected))
ureinfected$percentage<-(ureinfected$Freq/sum(ureinfected$Freq))*100
ureinfected
```

Out of 7817 observations only 3 cases were of reinfected. This is realistic on first year the reinfection was unheard of.

## Analyzing type variable
13. Show the number and percentage of "type" variable and discuss the result carefully with special reference to the missing values
```{r}
unique(covidtbl$type)
# Replacing the Missing value with Unknown
covidtbl$type[is.na(covidtbl$type)] <- 'unknown'
utype<-as.data.frame(table(covidtbl$type))
utype$percentage<-(utype$Freq/sum(utype$Freq))*100
utype
```
In this variable, we have more than 99 percentage of data as missing(NA). This means that we were not able to track where the people got covid from.

## Analyzing Nationality variable
14. Show the number and percentage of "nationality" variable and discuss the result carefully with special reference to the missing values
```{r}
unique(covidtbl$nationality)
# Replacing the Missing value with Unknown
covidtbl$nationality[is.na(covidtbl$nationality)] <- 'unknown'
unationality<-as.data.frame(table(covidtbl$nationality))
unationality$percentage<-(unationality$Freq/sum(unationality$Freq))*100
unationality
```

In this variable more than 99 percent of data was missing which we replaced with "Unknown".

## Cross Tabulation of province and current status
15. Show cross-tabulation of province (row variable) and current status (column variable) with row percentage and interpret the result carefully (table and prop.table can be used!)

## Cross Tabulation of sex and current status
16. Show the cross-tabulation of sex (row variable) and current status (column variable) with row percentage and interpret the result carefully

## cross-tabulation of age groups (row variable) and current status
17. Show the cross-tabulation of broad age groups (row variable) and current status (column variable) with row percentage and interpret the result carefully

## scatterplot of province (x-axis) and cleaned age (y-axis)
18. Show the scatterplot of province (x-axis) and cleaned age (y-axis) and get the appropriate correlation coefficient for this bi-variate data and interpret the result carefully
```{r}
plot(covidtbl$province,covidtbl$age, main = "Scatter Plot of province and age", xlab = "Province", ylab = "Age")
```

From the scatterplot above we can see that in Province 2, 3 and 4 older population got infected. In all the provience people from 15 to 60 got more infected.
## Show the scatterplot of age and diff1 
19. Show the scatterplot of age (x-axis) and diff1 (y-axis) and get the appropriate correlation coefficient for this bi-variate data and interpret the result carefully

```{r}
plot(covidtbl$age,covidtbl$diff1, main = "Scatter Plot of Age and Diff1", xlab = "Age", ylab = "Diff1")
```
From the scatter plot we can see that the number of days between recovered and reported dates
was higher for the age group around 35 to 40. This is to be expected as the average age was around 35.

## SWOT Analysis)
20. Summarize your learning using SWOT analysis to complete this project in a 2x2 table; be honest and specific so that I can learn from you too