---
title: "Assignmet4_1"
author: "Dipesh Poudel"
date: "10/5/2021"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web Scrapping. Getting data from Table in a website 

### Getting the Data from Website
```{r}
library(rvest)
library(dplyr)
wiki_link <- "https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Nepal"
# Reading the whole website
wiki_page <- read_html(wiki_link)
# Getting only the tables from the whole website
tables <- wiki_page %>% html_table(fill = TRUE)
# Printing a sample table
print(tables[[2]])
# Getting the Required Table
covid_table <- wiki_page %>%html_nodes("table") %>% .[16] %>%html_table() %>% .[[1]]
print(head(covid_table))
```
### Getting the Structure of the dataframe
```{r}
str(covid_table)
```

As we can see all the variables (columns) have `char` data type and also the first row and second row are heading. We need to resolve this. This process of cleaning up data is called scrubbing or Tiding the data.

### Removing the first row and last column
```{r}
#Changing column names: Column Underscore Row 1
names(covid_table) = paste(names(covid_table), covid_table[1, ], sep = "_")
#Removing first row
covid_table = covid_table[-1, ]
#Removing last column
covid_table <- covid_table[,-14]
#Viewing the data
View(covid_table)
#Checking structure of the data
str(covid_table)
```
### Renaming the columns

```{r}
#Renaming the column names with underscore between spaces
colnames(covid_table)
#Checking column names to do correct coding below
names(covid_table)[names(covid_table) == "Date_Date"] = "Date"
names(covid_table)[names(covid_table) == "Confirmed cases_Total"] = "Confirmed_Cases_Total"
names(covid_table)[names(covid_table) == "Confirmed cases_New"] = "Confirmed_Cases_New"
names(covid_table)[names(covid_table) == "Confirmed cases_Active"] = "Confirmed_Cases_Active"
names(covid_table)[names(covid_table) == "RT-PCR tests_Total"] = "PCR_Total"
names(covid_table)[names(covid_table) == "RT-PCR tests_New"] = "PCR_New"
names(covid_table)[names(covid_table) == "TPR_TPR"] = "TPR"
names(covid_table)[names(covid_table) == "RR_RR"] = "RR"
names(covid_table)[names(covid_table) == "CFR_CFR"] = "CFR"
str(covid_table)
```

In the data we can see that there are "+" and "%" sign. We need to remove those before analysis.

```{r}
#Removing + from four variables
covid_table$Confirmed_Cases_New = gsub('[+]', '',
covid_table$Confirmed_Cases_New)
covid_table$Recoveries_New = gsub('[+]', '', covid_table$Recoveries_New)
covid_table$Deaths_New = gsub('[+]', '', covid_table$Deaths_New)
covid_table$PCR_New = gsub('[+]', '', covid_table$PCR_New)
#Removing % from three variables
covid_table$TPR = gsub('[%]', '', covid_table$TPR)
covid_table$RR = gsub('[%]', '', covid_table$RR)
covid_table$CFR = gsub('[%]', '', covid_table$CFR)
str(covid_table)
```

All the variables have `char` as data type. We need to convert the data type to numeric or integer so that we can perform analysis.

### Converting chr variables as numbers and integers

```{r}
library(tidyverse)
covid_table<-covid_table%>%mutate_at(vars(Confirmed_Cases_Total,Confirmed_Cases_New,Confirmed_Cases_Active,Recoveries_Total,Recoveries_New,Deaths_Total,Deaths_New,PCR_Total,PCR_New),as.integer)
covid_table<-covid_table%>%mutate_at(vars(TPR,RR,CFR),as.numeric)
str(covid_table)
```
While doing the type casting (conversion) the places where there was no value it is replaced by `NA`.


```{r}
# Creating a sequence of date from 2020 Jan 23
date2<-seq(as.Date('2020-1-23'), by='days', length.out = 495)
# Adding the sequence of date to the dataframe
covid_table<-cbind(covid_table, date2)
```

```{r}
#Plot
plot(covid_table$date2, covid_table$Deaths_New,main = "Daily New Deaths",xlab = "Date",ylab = "Daily New Deaths")

plot(covid_table$date2, covid_table$Deaths_New, ylim =range(0:250),main = "Daily New Deaths",xlab = "Date",ylab = "Daily New Deaths")
```

As we can see in the first graph, there are three data points that are out of ordinary. For now we will replace these data points using data from other sources.

```{r}
#Repalce Deaths_New of 24 Feb as 1 in the data
covid_table[covid_table$date2=="2021-02-24", "Deaths_New"] = 1
covid_table[covid_table$date2=="2021-02-26", "Deaths_New"] = 4
covid_table[covid_table$date2=="2021-03-05", "Deaths_New"] = 0

plot(covid_table$date2, covid_table$Deaths_New, main = "Daily New Deaths",xlab = "Date",ylab = "Daily New Deaths")
```


```{r}
summary(covid_table$Deaths_New)
```

From summary above we can see that the maximum daily new deaths between 2020-01-23 and 2021-05-31 from covid in Nepal was 246. The mean median and the quartile values are low which means that the new deaths was relatively low for a long period of time and it exploded in the later time period.

```{r}
summary(covid_table$Deaths_Total)
```
The total number of deaths in Nepal was 984. However, this does not pain the accurate picture since the Deaths_Total column has null value from date 2020-11-02.  

```{r}
summary(covid_table$Deaths_CFR)
```
We got this result because we do not have column name Deaths_CFR. The correct column name is CFR.

```{r}
summary(covid_table$CFR)
```

This shows that the case fatality rate went up to 1.32%. This means that the at max 1.32% of total people who had covid died in Nepal from 2020-01-23 to 2021-05-31.

## Data Modeling
```{r}
#Simple Moving Average fit for new deaths data!
library(smooth)
sma = sma(covid_table$Deaths_New, h=14, silent=FALSE)
summary(sma)
forecast(sma)
plot(forecast(sma), main = "SMA(2) Forecast", ylab="Daily deaths")
```

## Importing Data From a Web API

```{r}
library(jsonlite)
url <- "https://data.askbhunte.com/api/v1/covid/timeline"
covidtbl <- fromJSON(txt=url, flatten=TRUE)
colnames(covidtbl)
summary(covidtbl)
```

