---
title: "Project 3"
author: "Dipesh Poudel"
date: "12/2/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project 3 - Data Visualization in R

## Part 1: Data visualization with Base R graphics packages 

Use the built-in CO2 data and do as follows: 


```{r}
# Loading the CO2 data
co2_data<-CO2
str(co2_data)
```
1. Create bar graph of plant, type and treatment variables
```{r}
variables<-c('Plant','Type','Treatment')
for (var in variables){
barplot(table(co2_data[var]),main = paste("Bar Graph of ",var),
        xlab = var, ylab = "Frequency")  
}

```
2. Create histogram of conc and uptake variables 
  Histogram of conc variable
```{r}
# Histogram of conc variable
hist(co2_data$conc,main = "Histogram of conc",xlab = "conc",ylab = "Frequency")
```
  Histogram of uptake variable
```{r}
# Histogram of uptake variable
hist(co2_data$uptake,main = "Histogram of uptake",xlab = "conc",ylab = "Frequency")
```

3. Create scatterplot of conc and uptake variables 

```{r}
plot(co2_data$conc,co2_data$uptake,main = "Scatter Plot of conc and uptake",
     xlab = "conc",ylab = "uptake")
```
4. Which measure of association is suitable for conc and uptake variables 
  Since the relationship is not linear we have to use spearman correlation for 
  association
5. Compute the best correlation coefficient for conc and uptake variables and interpret the result carefully.

```{r}
cor(co2_data$conc,co2_data$uptake,method  = c("spearman"))
```
Since the correlation coefficient is positive and greater than 0 we can say that the as the conc increases the uptake tends to increase but not in linear way.


## Part 2: Data visualization with ggplot2 package 
Use the built-in CO2 data and do as follows: 
```{r}
library(ggplot2)
```

1. Create bar graph of plant, type and treatment variables 

```{r}
variables<-c('Plant','Type','Treatment')
ggplot(data = co2_data) + geom_bar(mapping = aes(x = Plant))+ggtitle("Bar Graph of Plant")
ggplot(data = co2_data) + geom_bar(mapping = aes(x = Type))+ggtitle("Bar Graph of Type")
ggplot(data = co2_data) + geom_bar(mapping = aes(x = Treatment))+ggtitle("Bar Graph of Treatment")
```
2. Create histogram of conc and uptake variables 
```{r}
ggplot(data = co2_data)+geom_histogram(mapping = aes(x=conc))+ggtitle("Histogram of conc")
ggplot(data = co2_data)+geom_histogram(mapping = aes(x=uptake))+ggtitle("Histogram of uptake")
```

3. Create scatterplot of conc and uptake variables 
```{r}
ggplot(data=co2_data)+geom_point(mapping = aes(x=conc,y=uptake))+ggtitle('Scatter Plot of conc and uptake')
```

## Part 4: Text analysis with base/ggplot and Social Network Analysis with igraph package 
Use/load the attached “termDocMatrix.rdata” file in R studio and do as follows:
```{r}
file_path = 'data/termDocMatrix.rdata'
term_matrix_data<-load(file = file_path)
```

1. Covert this data as matrix 

```{r}
library(tm)
term_matrix_data<-as.DocumentTermMatrix(termDocMatrix,weighting=weightBin)
term_matrix_data<-as.matrix(term_matrix_data)
```

2. Get the term frequencies 
```{r}
freq <- sort(rowSums(term_matrix_data), decreasing=T)
freq
```

3. Create the histogram of the term frequencies 

```{r}
barplot(freq,main = "Frequency of Each Term",xlab = "Term",ylab = "Frequency")

hist(freq,main = "Histogram of Frequency",xlab = "Term Frequency")
```
4. Create the histogram of the terms with frequencies of 5 and more 
```{r}
freq_1<-subset(freq,freq>=5)
barplot(freq_1,main = "Frequency of Each Term with Freq>=5",xlab = "Term",ylab = "Frequency")

hist(freq-1,main = "Histogram of Frequency",xlab = "Term Frequency")
```
5. Create word cloud of the term frequencies 

```{r}
library(wordcloud)
freq <- sort(rowSums(term_matrix_data), decreasing=T)
wordcloud(words=names(freq), freq=freq, min.freq=5,
random.order=F)
```
6. Perform social network analysis of the termDocumentMatrix data and interpret it carefully  


```{r}
library(igraph)
#Transform Data into an Adjacency Matrix
termDocMatrix[termDocMatrix>=1] <- 1
# Transformation into term-term adjacency matrix
termMatrix <- termDocMatrix %*% t(termDocMatrix)
# Checking few terms in the adjacency matrix
termMatrix[1:5,1:5]
# Creating a undirected graph
g <- graph.adjacency(termMatrix, weighted=T, mode = "undirected")
# Removing the loop in same term 
g<-simplify(g)
plot(g)
```

In the graph above, we can see that the terms like 'r', 'mining', 'data' are at center and are frequently with other words. We can also see that 'time','series','introduction' have a cluster. We can also see the cluster of words 'research', 'postdoctoral','positions' from cluster. It makes sense for these words to come together.