---
title: "Assignment 8"
author: "Dipesh Poudel"
date: "1/5/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 8

**Use the built-in “mtcars” data of R and do as follows: **
1. Check the data with head(mtcars) and save a new data as mtcars.subset after dropping two non-numeric (binary) variables for PCA analysis 
```{r}
head(mtcars)
```


Removing the column with binary values

```{r}
mtcars.subset<-subset(mtcars,select = -c(vs,am))
```

2.Fit PCA in the as mtcars.pca matcars.subset data with cor = TRUE and scores = TRUE) 

```{r}
mtcars.pca<-princomp(mtcars.subset,cor=TRUE,scores = TRUE)
```
3. Get summary of mtcars.pca and interpret standard deviation, proportion of variance carefully 

```{r}
summary(mtcars.pca)
```
4. Get eigenvalue of the components using standard deviation of mtcars.pca and chose the number of components based on Kaiser’s criteria 
```{r}
# Eigenvalues are square of standard deviation
(eigenvalues<-mtcars.pca$sdev^2)
```
According to the Kaiser’s rule, PC with Eigenvalue >= 1 (SD-square)
must be used/retained for the latent variable. So, PC1 and PC1 are two components that we should retain.

5. Get scree plot and chose the number of components best on “first bend” of this plot 

```{r}
library(ggplot2)
# calculate total variance explained by each principal component
var_explained<-mtcars.pca$sdev^2/sum(mtcars.pca$sdev^2)

# Scree Plot

qplot(c(1:9),var_explained)+geom_line()+xlab("Principal Component")+ylab("Variance Explained")+ggtitle("Scree Plot")+ylim(0,1)
```

In the plot, the first bend is at 2 so 2 is the number of components to choose.

6. Write how many components must be retained based on Kaiser’s rule and/or scree plot 
Based on the scree plot and Kaiser's rule we have to retain two(2) components.

7. Fit the final PCA model based on the retained components and interpret it carefully 
```{r}
library(psych)
final_pca<-psych::principal(mtcars.subset,nfactors = 2,rotate = "none")
```

```{r}
final_pca
```

The first variable can explain 63% and the second variable can explain 23% variance. These both can explain 86% of the variance. 

8. Get the head of the saved loadings of mtcars.pca and interpret the values carefully 

```{r}
head(final_pca$loadings)
```

The `loadings` variable provides the coefficients of linear combination to compute pc1 and pc2 from the data. We can see that the cyl, disp,hp and wt 

9. Retain two components, get their loadings and interpret them carefully 

```{r}
final_pca$loadings
```

10. Get biplot of these two component loadings and interpret it carefully 

```{r}
biplot(final_pca,labels=rownames(mtcars.subset))
```

11. Get the head of the saved scores of mtcars.pca and interpret carefully 
```{r}
head(mtcars.pca$scores)
```

12 .Get the head of the scores of first two components of mtcars.pca and intepret it carefully 
```{r}
head(final_pca$scores)
```

13. Get biplot of these two component scores and interpret it carefully 
```{r}
biplot(final_pca,rownames(mtcars.subset))
```

14. Get dissimilar distance of all the variables of mtcars data as mtcars.dist 
```{r}
mtcars.dist<-dist(mtcars.subset)
```

15. Fit classical multi-dimensional scaling model with the mtcars.dist in 2-dimensional state as cars.mds.2d 

```{r}
cars.mds.2d<-cmdscale(mtcars.dist)
```

```{r}
summary(cars.mds.2d)
```

16. Plot the cars.mds.2d and compare it with the biplot of mtcars.pca and interpret it carefully 
```{r}
plot(cars.mds.2d,pch=19)
abline(h=0,v=0)
text(cars.mds.2d, pos = 4, labels =rownames(mtcars.subset), col ='tomato')
```

17. Fit classical multi-dimensional scaling model with the mtcars.dist in 3-dimensional state as cars.mds.3d 

```{r}
cars.mds.3d<-cmdscale(mtcars.dist,k=3)
```

```{r}
summary(cars.mds.3d)
```


18. Create a 3-d scatterplot of cars.mds.3d with type = “h”, pch=20 and lty.hplot=2 and interpret it carefully 

19. Create a 3-d scatterplot of cars.mds.3d with type = “h”, pch=20, lty.hplot=2 and color=mtcars$cyl and interpret it carefully 

20. Write a summary comparing PCA and MDS fits done above for mtcars data     