---
title: "13Dec"
author: "Dipesh Poudel"
date: "12/13/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 way ANOVA
```{r}
str(mtcars)
```

```{r}
mtcars$gear<-as.factor(mtcars$gear)
```

```{r}
str(mtcars$gear)
```

```{r}
with(mtcars,shapiro.test(mpg[gear==3]))
```
```{r}
with(mtcars,shapiro.test(mpg[gear==4]))
```
```{r}
with(mtcars,shapiro.test(mpg[gear==5]))
```
```{r}
library(car)
leveneTest(mpg~gear,data = mtcars)
```

```{r}
summary(aov(mpg~gear,data = mtcars))
```

```{r}
aov(mpg~gear,data = mtcars)
```

```{r}
TukeyHSD(aov(mpg~gear,data = mtcars))
```
First category is taken as reference.
```{r}
summary(lm(mpg~gear,data = mtcars))
```
In ANOVA we must adjust p-value in linear regression we do not perform adjustment

```{r}
pairwise.t.test(mtcars$mpg,mtcars$gear,p.adj="none")
```


Covariance

```{r}
cov(mtcars$wt,mtcars$mpg)
```

```{r}
cov(mtcars$wt,mtcars$mpg)
```

```{r}
mtcars$wt2<-mtcars$wt/2.2
```

```{r}
cov(mtcars$wt2,mtcars$mpg)
```

Correlation
```{r}
cor(mtcars$wt,mtcars$mpg)
```

```{r}
cor(mtcars$wt2,mtcars$mpg)
```

* LOW Degree:<0.25
* Medium Degree: 0.25-0.75
* High Degree: >0.75

Test the Correlation 
```{r}
cor.test(mtcars$wt,mtcars$mpg)
```

```{r}
cor.test(mtcars$wt2,mtcars$mpg)
```

Correlation + Test = Inference
Correlation = Descriptive

```{r}
lm1<-lm(mpg~wt,data = mtcars)
lm1
```
```{r}
summary(lm1)
```
LINE Test of Residuals
• L = Linearity of residuals
• I = Independence of residuals
• N = Normality of residuals
• E = Equal variance of residuals
```{r}
names(lm1)
```

```{r}
lm1.resid <- lm1$residuals
lm1.fitted <- fitted(lm1)
```

Linearity of residuals
```{r}
plot(lm1, which=1, col=c("blue"))
```
```{r}
summary(lm1$residuals)
```
Autocorrelation Function Plot (ACF) of the residuals
```{r}
acf(lm1$residuals)
```

```{r}
library(car)
durbinWatsonTest(lm1)
```

Normality of residuals
```{r}
plot(lm1, which=2, col=c("blue"))
```
```{r}
shapiro.test(lm1$residuals)
```

```{r}
plot(lm1, which=3, col=c("blue"))
```

```{r}
library(lmtest)
bptest(lm1)
```
Making the Prediction
```{r}
p <- data.frame(x=c(6))
colnames(p) <- "wt"
```

```{r}
predict(lm1, p)
```

