---
title: "Assignment 5"
author: "Dipesh Poudel"
date: "12/8/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 5

```{r cars}
#Define mean and SD
pop_mean<-50
pop_sd<-5
#Define lower and upper limits
LL <-pop_mean-pop_sd
UL <-pop_mean+ pop_sd
#Create a sequence of 100 x values based on pop mean and sd
x <-seq(-4,4, length=100)*pop_sd+pop_mean
y <-dnorm(x, pop_mean, pop_sd)
plot(x,y, type="l", lwd=2, axes=F, xlab="", ylab="")
sd_axis_bounds= 5
axis_bounds<-seq(-sd_axis_bounds*pop_sd + pop_mean, sd_axis_bounds*pop_sd+ pop_mean, by=pop_sd)
axis(side=1, at=axis_bounds, pos=0)
abline(v=50)
abline(v=50+pop_sd)
abline(v=50-pop_sd)
abline(v=50+2*pop_sd)
abline(v=50-2*pop_sd)
abline(v=50+3*pop_sd)
abline(v=50-3*pop_sd)
```
```{r}
library(ggplot2)
ggplot(data=as.data.frame(x,y),aes(x=x,y=y))+geom_line()+scale_x_continuous(breaks = axis_bounds)+geom_vline(xintercept = seq(30,70,by=pop_sd))+annotate(geom = "text",x = 50,y=mean(y)+0.01, label = "67% data")+annotate(geom = "segment",x = 45,y=mean(y)+0.012,xend = 55,yend = mean(y)+0.012)+annotate(geom = "text",x = 50,y=mean(y)+0.02, label = "95% data")+annotate(geom = "segment",x = 40,y=mean(y)+0.025,xend = 60,yend = mean(y)+0.025)+annotate(geom = "text",x = 50,y=mean(y)+0.03, label = "99% data")+annotate(geom = "segment",x = 35,y=mean(y)+0.035,xend = 65,yend = mean(y)+0.035)
```

```{r}
cars_data<-mtcars
ggplot(data=cars_data,aes(x=mpg))+geom_histogram(binwidth = 5)
```

```{r}
ggplot(data = cars_data,aes(mpg))+geom_histogram(binwidth = 5)+coord_flip()
```

```{r}
cars_data<-mtcars
qqnorm(cars_data$mpg, pch = 1, frame = FALSE)
qqline(cars_data$mpg, col = "steelblue", lwd = 2)
```
```{r}
shapiro.test(cars_data$mpg)
```
Null Hypothesis(H0): The data follows normal distribution
Alternative Hypothesis(H1): The data does not follow normal distribution
Since the p-value obtained from the test is not less than 0.05 we can say that the data comes from normal distribution i.e Null Hypothesis is accepted. 

Since the number of samples (32)<50 we can not use Kolmogorov-Smirnov. Thus we have to use 	Shapiro-Wilk normality test.
